### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@9fbaf07f388a6fd032325943e161934ee12c9d73 🚀

It is noteworthy that outside of South America, viral evolutionary
distinctiveness does not accurately tracks host diversity, with some
areas having over-distinct viruses (southern China but, oddly, not the
rest of southeast Asia). There are a number of likely explanations.
First, given the richness of bats in southeast Asia, many
betacoronaviruses likely remain to be discovered in this region. Indeed,
global predictions by (\textbf{Becker?}) highlight that southeast Asia
is a likely hostpot of unconfirmed hosts of betacoronaviruses, which
would likely result in additiona viral discoveries. This idea is
unsurprising given the growing realization, around the emergence of
SARS-CoV-2, that a unique lineage of similar viruses are widespread in
bats but still mostly undescribed. The most distinct
bats/betacoronavirus complex is found in South America, a region with a
comparatively lower number of hosts; this matches with the isolation
through viariance of the host group, and may highlight a different
co-evolutionary dynamic. Alternatively, this distinctiveness hostpot may
be a product of under-sampling: South-America is one of the places where
the fewest betacoronaviruses have been discovered (Anthony et al. 2017,
Olival et al. 2017, Allen et al. 2017), resulting in sparser
phylogenetic tree, thereby artificially inflating distinctiveness.
Adding more viruses would bring the distinctiveness of known sequences
down.

>>>>

It is noteworthy that outside of South America, viral evolutionary
distinctiveness does not accurately tracks host diversity, with some
areas having over-distinct viruses (southern China but, oddly, not the
rest of southeast Asia). There are a number of likely explanations.
First, given the richness of bats in southeast Asia, many
betacoronaviruses likely remain to be discovered in this region. Indeed,
global predictions by Becker et al. (2022) highlight that southeast Asia
is a likely hostpot of unconfirmed hosts of betacoronaviruses, which
would likely result in additiona viral discoveries. This idea is
unsurprising given the growing realization, around the emergence of
SARS-CoV-2, that a unique lineage of similar viruses are widespread in
bats but still mostly undescribed. The most distinct
bats/betacoronavirus complex is found in South America, a region with a
comparatively lower number of hosts; this matches with the isolation
through viariance of the host group, and may highlight a different
co-evolutionary dynamic. Alternatively, this distinctiveness hostpot may
be a product of under-sampling: South-America is one of the places where
the fewest betacoronaviruses have been discovered (Anthony et al. 2017,
Olival et al. 2017, Allen et al. 2017), resulting in sparser
phylogenetic tree, thereby artificially inflating distinctiveness.
Adding more viruses would bring the distinctiveness of known sequences
down.
